%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Let $N$ be a complete hyperbolic 3-manifold. $N$ can be realized as the
quotient manifold of a Kleinian group $\Gamma=\pi_1N$. This paper is devoted to
the following general question:

\begin{prob}

Does $\pi_1N$ contains any quasi-Fuchsian surface subgroups?

\end{prob}

In this paper, we answer this question in the special case where $N$ is a book
of $I$-bundles. We tackle this problem topologically, representing $N$ with
a pared 3-manifold $(M,P)$.  The pared structure $P\cin\bd M$ specifies the
locus corresponding to cusps of $N$. Finding a quasi-Fuchsian surface subgroup
of $\pi_1N$ amounts to asking whether we can find a closed immersed
$\pi_1$-injective surface in $M$ which avoids $P$ up to homotopy.  That is,
we're trying to choose a surface subgroup of $\pi_1M$ which trivially intersect
any of the peripherial subgroups associated to components of $P$. We prove full
results on which $(M,P)$ admit a surface subgroup in the book of $I$-bundles
case.

To motivate this choice of problem and special case to solve, we briefly
discuss the history of this problem and previous work on it. We begin with
a straightforward, purely topological question:

\begin{prob}

Which 3-manifold groups contain surface subgroups? Equivalently, which
3-manifolds contain closed $\pi_1$-injective surfaces of negative Euler
characteristic?

\end{prob}

This problem arises quite directly when considering 3-manifold groups. The
first explicit answer was proposed by Waldhausen \cite{Kirby} in
the Surface Subgroup Conjecture: every closed irreducible 3-manifold with
infinite $\pi_1$ contains a surface subgroup. As originally stated, the
Conjecture only applies to closed 3-manifolds. By geometrization, we can
consider two cases: Seifert fibered spaces or, more generally, graph manifolds,
and manifolds with at least one hyperbolic piece. For graph manifolds, the
problem was fully solved by Neumann \cite{Neu}.  After some preliminary
efforts in the hyperbolic case by Cooper--Long \cite{CooperLong} and Li
\cite{Li}, it was finally resolved in general by Kahn and Markovic in
2012 \cite{KM}.  In fact, the surface subgroups they find are quasi-Fuchsian,
and they also obtained a density result stating that there are many such
surfaces \cite{KM2}.

We can naturally ask this question for manifolds with boundary as well.
Cooper, Long, and Reid showed \cite{CLR} that a compact connected irreducible
3-manifold with non-empty incompressible boundary  must contain an essential
closed surface, unless it's covered by a product $F\times I$. In particular,
this shows that every non-closed hyperbolic 3-manifold group contains a surface
subgroup.

However, since the result of Kahn and Markovic provided us with surface
subgroups which are quasi-Fuchsian, we might ask whether we can do this in the
non-closed case as well. Note that the Cooper-Long-Reid result is insufficient,
as the surfaces obtained may contain accidental parabolics, that is, overlap in
$\pi_1$ with the pared locus $P$. If this occurs they will not be
quasi-Fuchsian. This does commonly occur in practice --- see other work of

Cooper--Long--Reid \cite{CLRbundles} and Menasco--Reid \cite{MenascoReid} for
examples.  Of course, in the closed case there cannot be any parabolics to
worry about, as the boundary is empty.

More recently, there have been efforts to find quasi-Fuchsian surface subgroups
specifically. Masters and Zhang \cite{MZ} showed that every hyperbolic knot
complement contains a quasi-Fuchsian surface subgroup, and later extended this
result to link complements \cite{MZ2}. This fully addresses the case of finite
volume cusped hyperbolic 3-manifolds. An alternate proof was provided by Baker
and Cooper \cite{BC}. The densite result of Kahn--Markovic was also extended by
work of Cooper and Futer \cite{CooperFuter}.

The remaining case is infinite volume hyperbolic 3-manifolds. This paper
presents a preliminary approach to the problem. We outline a general argument
for all infinite volume hyperbolic 3-manifolds. We explicitly solve a special
case, books of $I$-bundles, which we believe will generalize. We prove that all
books of $I$-bundles admit quasi-Fuchsian surface subgroups, except for a few
obvious negative cases where the pared locus is too large. We now present
a brief outline of the topics we'll cover.

In Chapter 2, we provide some brief background on relevant topics. We define
quasi-Fuchsian subgroups and describe various criteria we can use to establish
that a given surface subgroup is quasi-Fuchsian. We define pared 3-manifolds
and explain how to translate these criteria to topological criteria for
surfaces in pared 3-manifolds. We briefly discuss the case of geometrically
infinite hyperbolic 3-manifolds. We restrict to the geometrically finite case,
with reference to hyperbolization.

In Chapter 3, we outline a general procedure to either find a closed
quasi-Fuchsian surface in a pared 3-manifold, or conclude that there is no such
surface.  We decompose the pared 3-manifold along disks and annuli in an
alternating hierarchy, and prove that any quasi-Fuchsian surface ond pared
structure behave nicely under this decomposition. We prove that the hierarchy
is finite and terminates in balls and acylindrical 3-manifolds. We briefly
describe strategies for the case of an acylindrical piece, and the case of no
acylindrical pieces. We explain how these might rely on the books of
$I$-bundles theorem, motivating our study of that special case.

In Chapter 4, we prove preliminary facts we will need for the main theorem on
books of $I$-bundles. The quasi-Fuchsian surface construction relies on
subgroup separability arguments to simplify the 3-manifold under consideration.
So we give a brief overview of subgroup separability, and prove key topological
lemmas using it that we'll need later. We also prove many topological
properties of books of $I$-bundles, and quasi-Fuchsian surfaces and pared loci
inside them.  These results refine the general decomposition properties proven
in Chapter 3.  We use these properties to prove a criterion that establishes
whether or not a given closed surface in a book of $I$-bundles is
quasi-Fuchsian.  Finally, we use this criterion to explicitly identify closed
quasi-Fuchsian surfaces in an example book of $I$-bundles. We describe which
pared structures on this example allow closed quasi-Fuchsian surfaces to be
present. This example will motivate our main theorem.

In Chapter 5, we state and prove the main theorem on books of $I$-bundles. We
first describe a reduction process that removes parts of the book of
$I$-bundles that cannot intersect a closed quasi-Fuchsian surface. We state and
prove the negative case of the main theorem --- that is, which pared books of
$I$-bundles do not admit quasi-Fuchsian surfaces.

We then state and prove the positive case, that all other pared books of
$I$-bundles do admit quasi-Fuchsian surfaces. This proof relies on an elaborate
construction. First, we use the separability properties from Chapter 4 to
simplify the topology of the book of $I$-bundles $M$ and its pared locus $P$.
We then construct a surface with boundary corresponding to each boundary
component of $M$, with certain nice properties with respect to $P$. We then
attach these surfaces with boundary together to form a closed surface. Our
construction allows us to use the criterion from Chapter 4 to show that the
resulting surface is quasi-Fuchsian. This will complete the main proof.

Note that all later chapters depend on Chapter 2. However, Chapters 4 and 5 may
be read independently of Chapter 3. Chapter 4 is largely elementary, and
readers familiar with subgroup separability and cut-and-paste arguments may
wish to read only the definitions and the statement of the pared lifting
criterion before moving to Chapter 5.
