%% thesis.tex 2014/04/11
%
% Based on sample files of unknown authorship.
%
% The Current Maintainer of this work is Paul Vojta.

\documentclass{ucbthesis}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{rotating}
\usepackage{url}
\usepackage{etoolbox}
%\usepackage{biblatex}
\usepackage{amsmath,amscd,amssymb,amsthm,amsfonts}

% To compile this file, run "latex thesis", then "biber thesis"
% (or "bibtex thesis", if the output from latex asks for that instead),
% and then "latex thesis" (without the quotes in each case).

% Double spacing, if you want it.  Do not use for the final copy.
% \def\dsp%{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}

\input{commands}

\newtoggle{paper}

% Choose whether or not to include the "paper pages" (blank pages to force
% certain pages to odd/even. I worked out where they go by hand, so it's
% a bit messy...
\togglefalse{paper}

\newcommand{\paperpage}{\iftoggle{paper}{\newpage\null\thispagestyle{empty}\newpage}{}}

\begin{document}

% Declarations for Front Matter

\title{Quasi-Fuchsian surface subgroups of infinite covolume Kleinian groups}
\author{Alvin Jay Kerber}
\degreesemester{Summer}
\degreeyear{2017}
\degree{Doctor of Philosophy}
\chair{Professor Ian Agol}
\othermembers{Professor Denis Auroux \\
  Professor Martin Olsson \\
  Professor Martin White}
\numberofmembers{4}
% Previous degrees are no longer to be listed on the title page.
% \prevdegrees{B.A. (University of Northern South Dakota at Hoople) 1978 \\
%   M.S. (Ed's School of Quantum Mechanics and Muffler Repair) 1989}
\field{Mathematics}
% Designated Emphasis -- this is optional, and rare
% \emphasis{Colloidal Telemetry}
% This is optional, and rare
% \jointinstitution{University of Western Maryland}
% This is optional
\campus{Berkeley}

% For a masters thesis, replace the above \documentclass line with
% \documentclass[masters]{ucbthesis}
% This affects the title and approval pages, which by default calls this
% document a "dissertation", not a "thesis".

\maketitle
% Delete (or comment out) the \approvalpage line for the final version.
%\approvalpage
\paperpage

\copyrightpage
\paperpage

\include{abstract}
\paperpage

\iftoggle{paper}{}{\begin{frontmatter}}

\iftoggle{paper}{\thispagestyle{empty}}{}

\begin{dedication}
\null\vfil
\begin{center}

To Mommy and Daddy

\end{center} \vfil\null \end{dedication}

\paperpage

\iftoggle{paper}{\begin{frontmatter}}{}

% You can delete the \clearpage lines if you don't want these to start on
% separate pages.

\tableofcontents
\clearpage
\listoffigures

%\clearpage
%\listoftables

\begin{acknowledgements}

First and foremost, thank you to my advisor, Ian Agol. Without your guidance,
leadership, and support, this dissertation would not exist. Thanks also to my
committee: Denis Auroux, Martin Olsson, and Martin White.

Thank you to Jason Manning and Saul Schleimer for valuable suggestions.  Thank
you to Kevin Donoghue, Kyle Miller, and Franco Vargas Pallete for many good
conversations, and lots of help and support.

Thank you to Richard Schwartz and Jeffrey Brock of Brown University, who first
taught me about geometry and topology. Thanks also to Zev Chonoles, Andrew
Furnas, Alex Kruckman, and Aaron Mazel-Gee for a great time at Brown doing
math. Your enthusiasm is contagious.

Thank you to Paul Bamberg, Emmanuel ``Manny'' Sodbinow, and Robert Winters for
getting me interested in advanced math.

Thank you to all my friends and colleagues in the math graduate program.
Thanks especially to Shishir Agrawal, Harrison Chen, Semyon Dyatlov, Julia
Erhard, Anna Lieb, Gus Schrader, Noah Schweber, Clare Stichcombe, Ethan Van
Andel, and Markus Vasquez.  I wouldn't have survived my first year in Berkeley
without you.

Thank you to all the staff of the math department. You've gone above and beyond
your responsibilities to help me out. Thanks especially to Barb Waller, who
kept me going during my hardest year.

My time at Berkeley would not have been the same without the Cal Ballroom dance
team and Team 106. Thank you to all my coaches and fellow dancers, who have
always been there for me.  My biggest thanks to Brittany Whitehill.  You've
been a great partner and friend.

Thank you to my all my friends and family. You kept me on track and pointed the
right way, and you were there for me when I needed you the most.

Finally, thank you to Katharine, for everything. You're the best.

\end{acknowledgements}

\paperpage

\end{frontmatter}

\pagestyle{headings}

\include{chap1}
\include{chap2}
\include{chap3}
\include{chap4}
\include{chap5}

% Add/remove depending on page count of main doc (awkward)
\paperpage

\bibliographystyle{plain}
\bibliography{refs}

\end{document}
